--- 
title: "Power Analysis with Superpower"
author: "Aaron R. Caldwell & Dani&euml;l Lakens"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib, references.bib]
biblio-style: apalike
link-citations: yes
description: "This is a book describing the capabilities of the Superpower R package."
---

This is a compilation of documents for [Superpower](https://github.com/arcaldwell49/Superpower) R package written in **Markdown** [@R-rmarkdown] and compiled by **Bookdown** [@R-bookdown].



```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 
  "pwr", "pwr2ppl",
  "multcomp"
), 'packages.bib')
```

# Preface {-}

The goal of `Superpower` is to easily simulate factorial designs and empirically calculate power using a simulation approach. 
The R package is intended to be utilized for prospective (a priori) power analysis. Please don't be silly and compute [post hoc power](https://discourse.datamethods.org/t/reference-collection-to-push-back-against-common-statistical-myths/1787). 

This package, and book, expect readers to have some familiarity with R [@R-base]. However, we have created two Shiny apps (for the [`ANOVA_power`](http://shiny.ieis.tue.nl/anova_power/) & [`ANOVA_exact`](http://shiny.ieis.tue.nl/anova_exact/) functions respectively) to help use these functions if you are not familiar with R.

In this book we will display a variety of ways the `Superpower` package can be used for power analysis and sample size planning for factorial experimental designs. We also included various examples of the performance of `Superpower` against other R packages (e.g., `pwr2ppl` by @R-pwr2ppl  and `pwr` by @R-pwr) and statistical programs (such as G*Power @faul2007g by  or MOREpower  @Campbell2012MorePower6F). All uses of the `ANOVA_power` function have been run with 10000 iterations (`nsims = 10000`). If you have any issues using `Superpower` or want to expand its capabilities please raise the issue on our [GitHub](https://github.com/arcaldwell49/Superpower/issues) repository.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#set number of simulations RESET BEFORE PUBLICATION TO 100.000
nsims <- 10000 
library(mvtnorm)
library(MASS)
library(afex)
library(emmeans)
library(gridExtra)
library(reshape2)
library(pwr)
library(pwr2ppl)
library(lsr)
library(viridis)
library(broom)
library(car)
library(tidyverse)
library(Superpower)
library(knitr)
library(kableExtra)

#latex_options = "hold_position"
```

```{r global_options, include=FALSE}
#options(knitr.table.format = "latex")

```

